---
title: "p8105_hw6_xz3173"
author: "Xue Zhang"
date: "2023-11-28"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Due date

Due: December 3 at 11:59pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |


### Problem 0

This "problem" focuses on structure of your assignment, including the use of R Markdown to write reproducible reports, the use of R Projects to organize your work, the use of relative paths to load data, and the naming structure for your files. 

To that end: 

* create a public GitHub repo + local R Project for this assignment
* write solutions using a .Rmd file that outputs a `github_document` / .md file
* submit a link to your repo via Courseworks

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems. 

For Problem 0, we will assess adherence to the instructions above regarding repo structure, git commit history, and whether we are able to knit your .Rmd to ensure that your work is reproducible. Adherence to appropriate styling and clarity of code will be assessed in Problems 1+ using the homework [style rubric](homework_style_rubric.html). 

This homework includes figures; the readability of your embedded plots (e.g. font sizes, axis labels, titles) will be assessed in Problems 1+.

```{r load_libraries}
library(tidyverse)
library(modelr)
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



### Problem 2

For this problem, we’ll use the Central Park weather data similar to data we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:

r̂ 2
log(β̂ 1∗β̂ 2)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2
 and log(β̂ 0∗β̂ 1)
. Note: broom::glance() is helpful for extracting r̂ 2
 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 1∗β̂ 2)
.


```{r}
set.seed(1)

# Bootstrap for r_squared
bootstrap_results =
  weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df) ),
    results = map(models, broom::glance)) |>
  select(-strap, -models) |>
  unnest(results)

# Pull r_squared
r_squared = 
  bootstrap_results |>
  pull(r.squared) 

# Plot of r_squared
bootstrap_results |>
  ggplot(aes(x = r_squared)) +
  geom_density()
```

```{r}
# Confidence interval for r_squared
bootstrap_results |>
  summarize(
    r_squared_ci_lower = quantile(r_squared, 0.025),
    r_squared_ci_upper = quantile(r_squared, 0.975))
```

```{r}
# Bootstrap for log_beta_products
bootstrap_results =
  weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df) ),
    results = map(models, broom::tidy)) |>
  select(-strap, -models) |>
  unnest(results)


# Pull beta_1
beta_1 =
  bootstrap_results |>
  filter(term == "tmin") |>
  pull(estimate) 

# Pull beta_2
beta_2 =
  bootstrap_results |>
  filter(term == "prcp") |>
  pull(estimate)

# Create a data frame from beta_1 and beta_2
beta_df = data.frame(beta_1, beta_2)

# Filter out rows where the product of beta_1 and beta_2 is not positive and compute log_beta_product
beta_df = beta_df |>
  filter(beta_1 * beta_2 > 0) |>
  mutate(
    log_beta_product = log(beta_1 * beta_2))

# Plot of log_beta_product
beta_df |>
  ggplot(aes(x = log_beta_product)) +
  geom_density()
```

```{r}
# Confidence Interval for log_beta_product
beta_df |>
  summarize(
    log_beta_product_ci_lower = quantile(log_beta_product, 0.025),
    log_beta_product_ci_upper = quantile(log_beta_product, 0.975))
```


### Problem 3

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

babysex: baby’s sex (male = 1, female = 2)
bhead: baby’s head circumference at birth (centimeters)
blength: baby’s length at birth (centimeteres)
bwt: baby’s birth weight (grams)
delwt: mother’s weight at delivery (pounds)
fincome: family monthly income (in hundreds, rounded)
frace: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
gaweeks: gestational age in weeks
malform: presence of malformations that could affect weight (0 = absent, 1 = present)
menarche: mother’s age at menarche (years)
mheigth: mother’s height (inches)
momage: mother’s age at delivery (years)
mrace: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
parity: number of live births prior to this pregnancy
pnumlbw: previous number of low birth weight babies
pnumgsa: number of prior small for gestational age babies
ppbmi: mother’s pre-pregnancy BMI
ppwt: mother’s pre-pregnancy weight (pounds)
smoken: average number of cigarettes smoked per day during pregnancy
wtgain: mother’s weight gain during pregnancy (pounds)

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
# Load data
birth_weight_df =
  read_csv("./data/birthweight.csv")

summary(birth_weight_df)
str(birth_weight_df)


# Convert Numeric to Factor
birth_weight_df |>
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    mrace = as.factor(mrace),
    malform = as.factor(malform)
  )

# Check for missing data
colSums(is.na(birth_weight_df))

# Check for negative or zero values in birth weight
sum(birth_weight_df$bwt <= 0)

# Similarly, check other variables where negative values don't make sense
sum(birth_weight_df$delwt <= 0) # Mother's weight at delivery
sum(birth_weight_df$gaweeks <= 0) # Gestational age in weeks


# Check for extremely high values in birth weight
sum(birth_weight_df$bwt > 5000) # You can set an appropriate threshold

# Check for other variables as well
sum(birth_weight_df$gaweeks > 42) # Gestational age greater than 42 weeks

sum(birth_weight_df$gaweeks < 37) 

# Boxplot for birth weight
boxplot(birth_weight_df$bwt, main="Boxplot of Birth Weight")
boxplot(birth_weight_df$bhead)
boxplot(birth_weight_df$blength)
boxplot(birth_weight_df$delwt)
boxplot(birth_weight_df$fincome)
boxplot(birth_weight_df$gaweeks)
boxplot(birth_weight_df$menarche)
boxplot(birth_weight_df$mheight)
boxplot(birth_weight_df$momage)
boxplot(birth_weight_df$ppbmi)
boxplot(birth_weight_df$ppwt)
boxplot(birth_weight_df$smoken)
boxplot(birth_weight_df$wtgain)


# Histograms for continuous variables
hist(birth_weight_df$gaweeks, main="Histogram of Gestational Weeks")

# Check unique values for categorical variables
unique(birth_weight_df$babysex)
unique(birth_weight_df$frace)
unique(birth_weight_df$mrace)
unique(birth_weight_df$malform)

# Check if any records have pre-pregnancy weight more than delivery weight
sum(birth_weight_df$ppwt > birth_weight_df$delwt)

```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

Step 1: Hypothesis and Initial Model Formulation
Based on medical literature and common understanding, several factors are known to influence birthweight. These include:

Gestational Age (gaweeks): Typically, longer gestation leads to higher birth weight.
Mother’s Health and Characteristics: Such as pre-pregnancy BMI (ppbmi), weight gain during pregnancy (wtgain), height (mheight), age (momage).
Baby’s Characteristics: Such as sex (babysex).
Parental Characteristics: Including family income (fincome), mother's and father's race (mrace, frace).
Lifestyle Factors: Smoking during pregnancy (smoken).

Step 2: Data-Driven Refinement
Using statistical techniques, refine your model by:

Adding Interaction Terms: Consider adding interaction terms if you hypothesize that the effect of one variable depends on another (e.g., the effect of smoken might differ by momage).
Checking Multicollinearity: Use Variance Inflation Factor (VIF) to check for multicollinearity among independent variables.
Model Selection Criteria: Utilize AIC, BIC, or adjusted R-squared to compare models and select the best fitting one.

Step 3: Model Fitting
Fit the linear regression model in R using the lm() function:

```{r}
model = lm(bwt ~ gaweeks + ppbmi + wtgain + mheight + momage + babysex + fincome + mrace + frace + smoken, data = birth_weight_df)

summary(model)

```

Step 4: Diagnostics and Residual Analysis
After fitting the model, it’s crucial to perform residual analysis to check the assumptions of linear regression:

Residual vs. Fitted Plot: To check for homoscedasticity and linearity.
Normal Q-Q Plot: To check if residuals are normally distributed.
Scale-Location Plot: To check homoscedasticity again.
Residuals vs Leverage Plot: To identify influential outliers.
In R, you can use the plot() function on the model object to generate these diagnostic plots.

Step 5: Show Plot of Model Residuals Against Fitted Values
You can use ggplot2 and broom packages in R for a more customized plot of residuals against fitted values.

```{r}
# Add predictions and residuals to the original dataset
augmented_df = birth_weight_df |>
  add_predictions(model) |>
  add_residuals(model)

# Plotting residuals against fitted values
ggplot(augmented_df, aes(x = pred, y = resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residuals vs Fitted Values")


```

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)

```{r}
# Model 2
model_2 = lm(bwt ~ blength + gaweeks, data = birth_weight_df)
```

One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
# Model 3
model_3 = lm(bwt ~ bhead * blength * babysex, data = birth_weight_df)
```

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

```{r}
# Define the number of folds for cross-validation
folds = crossv_mc(birth_weight_df, 10) # For 10-fold cross-validation

# Function to calculate RMSE
calc_rmse = function(model, data) {
  pred = predict(model, newdata = data)
  sqrt(mean((data$bwt - pred)^2))
}

# Applying cross-validation to each model
rmse_1 = map_dbl(folds$train, ~calc_rmse(lm(bwt ~ gaweeks + ppbmi + wtgain + mheight + momage + babysex + fincome + mrace + frace + smoken, data = .x), folds$test[[1]]))

rmse_2= map_dbl(folds$train, ~calc_rmse(lm(bwt ~ blength + gaweeks, data = .x), folds$test[[1]]))

rmse_3 = map_dbl(folds$train, ~calc_rmse(lm(bwt ~ bhead * blength * babysex, data = .x), folds$test[[1]]))

# Calculating the mean RMSE for each model
mean_rmse_1 = mean(rmse_1)
mean_rmse_2 = mean(rmse_2)
mean_rmse_3 = mean(rmse_3)

```