---
title: "p8105_hw6_xz3173"
author: "Xue Zhang"
date: "2023-11-28"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Due date

Due: December 3 at 11:59pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |


### Problem 0

This "problem" focuses on structure of your assignment, including the use of R Markdown to write reproducible reports, the use of R Projects to organize your work, the use of relative paths to load data, and the naming structure for your files. 

To that end: 

* create a public GitHub repo + local R Project for this assignment
* write solutions using a .Rmd file that outputs a `github_document` / .md file
* submit a link to your repo via Courseworks

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems. 

For Problem 0, we will assess adherence to the instructions above regarding repo structure, git commit history, and whether we are able to knit your .Rmd to ensure that your work is reproducible. Adherence to appropriate styling and clarity of code will be assessed in Problems 1+ using the homework [style rubric](homework_style_rubric.html). 

This homework includes figures; the readability of your embedded plots (e.g. font sizes, axis labels, titles) will be assessed in Problems 1+.

```{r load_libraries}
library(tidyverse)
library(modelr)
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



### Problem 2

For this problem, we’ll use the Central Park weather data similar to data we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:

r̂ 2
log(β̂ 1∗β̂ 2)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2
 and log(β̂ 0∗β̂ 1)
. Note: broom::glance() is helpful for extracting r̂ 2
 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 1∗β̂ 2)
.


```{r}
set.seed(1)

# Bootstrap for r_squared
bootstrap_results =
  weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df) ),
    results = map(models, broom::glance)) |>
  select(-strap, -models) |>
  unnest(results)

# Pull r_squared
r_squared = 
  bootstrap_results |>
  pull(r.squared) 

# Plot of r_squared
bootstrap_results |>
  ggplot(aes(x = r_squared)) +
  geom_density()
```

**The distribution of `r_squared` values from the 5000 bootstrap samples indicates a strong average model performance, with the mean `r_squared` around `r mean(r_squared)`, which indicates a better fit of the model to the data. This high mean value suggests that, on average, the model explains a large portion of the variance in the response variable `tmax` using predictors `tmin` and `prcp`. The distribution of these `r_squared` values is approximately across the bootstrap samples.**

```{r}
# Confidence interval for r_squared
bootstrap_results |>
  summarize(
    r_squared_ci_lower = quantile(r_squared, 0.025),
    r_squared_ci_upper = quantile(r_squared, 0.975)) |>
  knitr::kable(digits = 3)
```


**The 95% confidence interval for `r_squared` indicates that we can be 95% confident that the true coefficient of determination `r_squared` for the population lies between `0.889` and `0.941`. These high values of `r_squared` suggest that the model, with `tmin` and `prcp` as predictors, does a good job of explaining the variability in `tmax`.**

```{r}
# Bootstrap for log_beta_products
bootstrap_results =
  weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df) ),
    results = map(models, broom::tidy)) |>
  select(-strap, -models) |>
  unnest(results)


# Pull beta_1
beta_1 =
  bootstrap_results |>
  filter(term == "tmin") |>
  pull(estimate) 

# Pull beta_2
beta_2 =
  bootstrap_results |>
  filter(term == "prcp") |>
  pull(estimate)

# Create a data frame from beta_1 and beta_2
beta_df = data.frame(beta_1, beta_2)

# Filter out rows where the product of beta_1 and beta_2 is not positive and compute log_beta_product
beta_df = beta_df |>
  filter(beta_1 * beta_2 > 0) |>
  mutate(
    log_beta_product = log(beta_1 * beta_2))

# Plot of log_beta_product
beta_df |>
  ggplot(aes(x = log_beta_product)) +
  geom_density()
```

**The distribution of `log_beta_products` from the 5000 bootstrap samples, after filtering out instances where the product of `beta_1` and `beta_2` is non-positive, shows a left-skewed (or negatively skewed) pattern. This skewness is characterized by a longer tail extending towards lower values, with the majority of the data points and the median skewed towards the right. The median of the distribution is approximately `r median(beta_df$log_beta_product)`, which suggests that, while most bootstrap samples indicate a certain level of combined effect of `tmin` and `prcp` on `tmax`, there are a notable number of samples where this combined effect is smaller. This left skewness can imply that in several instances, the interaction or combined influence of `tmin` and `prcp` is less pronounced than what is observed on average in the dataset.**

**It's important to consider that the filtering criteria may have excluded cases where the product of `beta_1` and `beta_2` is non-positive, may have impacted the analysis. This exclusion potentially omits scenarios where the predictors have an inverse or negligible combined effect on the response variable, possibly leading to a skewed representation of their interaction.**

**In the context of this analysis, the observed left skewness in the distribution of `log_beta_products` suggests that while the combined effect of `tmin` and `prcp` generally contributes positively to `tmax`, the strength of this relationship varies, with a noteworthy number of instances exhibiting a less-than-average effect. This variation underscores the complexity of the relationship between these variables and highlights the need to consider a range of scenarios when interpreting the impact of `tmin` and `prcp` on `tmax`.**



```{r}
# Confidence Interval for log_beta_product
beta_df |>
  summarize(
    log_beta_product_ci_lower = quantile(log_beta_product, 0.025),
    log_beta_product_ci_upper = quantile(log_beta_product, 0.975)) |>
  knitr::kable(digits = 3)
```


**The 95% confidence interval for `log_beta_products` suggests that we can be 95% confident that the true value of `log_beta_products` lies between`-8.885`and `-4.604`in the population from the sample. The negative values here indicate that the product of `beta_1` and `beta_2` is typically less than1, as the logarithm of a number less than 1 is negative. This might suggest a less pronounced or even an inverse relationship between the combined effect of the predictors and the response variable.**



### Problem 3

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

babysex: baby’s sex (male = 1, female = 2)
bhead: baby’s head circumference at birth (centimeters)
blength: baby’s length at birth (centimeteres)
bwt: baby’s birth weight (grams)
delwt: mother’s weight at delivery (pounds)
fincome: family monthly income (in hundreds, rounded)
frace: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
gaweeks: gestational age in weeks
malform: presence of malformations that could affect weight (0 = absent, 1 = present)
menarche: mother’s age at menarche (years)
mheigth: mother’s height (inches)
momage: mother’s age at delivery (years)
mrace: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
parity: number of live births prior to this pregnancy
pnumlbw: previous number of low birth weight babies
pnumgsa: number of prior small for gestational age babies
ppbmi: mother’s pre-pregnancy BMI
ppwt: mother’s pre-pregnancy weight (pounds)
smoken: average number of cigarettes smoked per day during pregnancy
wtgain: mother’s weight gain during pregnancy (pounds)

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
# Load data
birth_weight_df =
  read_csv("./data/birthweight.csv")

# Convert Numeric to Factor
birth_weight_df |>
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    mrace = as.factor(mrace),
    malform = as.factor(malform)
  )

# Check for missing data
colSums(is.na(birth_weight_df))
```


Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
# Example of starting with a simple linear regression model
model_1 = lm(bwt ~ blength + delwt + gaweeks + ppbmi + smoken, data = birth_weight_df)
```

```{r}
# Add predictions and residuals
model_1_df = birth_weight_df |>
  add_predictions(model_1) |>
  add_residuals(model_1) |>

# Plot of residuals against fitted value
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") +
  facet_grid(~model)
  labs(x = "Fitted values", y = "Residuals", title = "Residuals vs. Fitted Plot")
```



Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)

```{r}
model_2 = lm(bwt ~ blength + gaweeks, data = birth_weight_df)
```


One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
model_3 = lm(bwt ~ bhead * blength * babysex, data = birth_weight_df)
```


Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

```{r}
# Cross-validation for each model
cv_df = 
  crossv_mc(birth_weight_df, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

```{r}
cv_df = 
  cv_df |>
  mutate(
    model_1 = map(train, \(df) lm(bwt ~ blength + delwt + gaweeks + ppbmi + smoken, data = df)),
    model_2 = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model_3 = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))) |>
  
  # Compute RMSE for each model
  mutate(
    rmse_model_1 = map2_dbl(model_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model_2 = map2_dbl(model_2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model_3 = map2_dbl(model_3, test, \(mod, df) rmse(model = mod, data = df)))
```

```{r}
cv_df |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") |>
  mutate(model = fct_inorder(model)) |>
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```
